% !TEX root = ../vr_st.tex

\subsection{Sums and products}

\subsubsection{}

For any two pointed $\R$-spaces $X$ and $Y$, the pointed $\R$-space $X \vee Y$, their \defn{wedge sum}, is defined by
\[
(X \vee Y)_r = X_r \vee Y_r \quad\text{and}\quad (X \vee Y)_{s,t} = (X_{s,t} \vee Y_{s,t}).
\]

\medskip\theorem
Let $X$ and $Y$ be pointed cellular $\R$-spaces.
For any linear cohomology operation $\theta$:
\begin{align*}
	\img_\theta(X \vee Y) &\cong \img_\theta(X) \oplus \img_\theta(Y), \\
	\ker_\theta(X \vee Y) &\cong \ker_\theta(X) \oplus \ker_\theta(Y).
\end{align*}

\begin{proof}
	For any \(r \in \R\), the inclusions $X_r \to X_r \vee Y_r \leftarrow Y_r$ induce an isomorphism $\rH^*(X_r \vee Y_r) \to \rH^*(X_r) \oplus \rH^*(Y_r)$ (\cite[Corollary 2.25.]{hatcher2000}).
	The claim follows from the naturality of \(\theta\).\anibal{note to self: Say more?}
%	By the naturality of $\theta$ we have that $\theta_{X \vee Y} \circ (\iota^*_X \oplus \iota^*_Y) = (\iota^*_X \circ \theta_X) \oplus (\iota^*_Y \circ \theta_Y)$.
%	So, $\img\theta_{X \vee Y} \cong \img\theta_X \oplus \img\theta_Y$ and $\ker\theta_{X \vee Y} \cong \ker\theta_X \oplus \ker\theta_Y$.
\end{proof}

\subsubsection{}\label{ss:wedge sum}

The \defn{gluing metric} on the wedge sum of pointed metric spaces ${(\cX_i, x_i)}_{i = 0}^k$ is defined as follows: For $x \in \cX_i$ and $x' \in \cX_j$, let
\[
d(x, x') =
\begin{cases}
	d_{\cX_i}(x, x_i) + d_{\cX_j}(x', x_j) & i \neq j, \\
	\hfil d_{\cX_i}(x, x') & i = j.
\end{cases}
\]
%if $i \neq j$, and
%\[
%d(x, x') = d_{\cX_i}(x, x')
%\]
%if $i =j$,

Recall from \cite[Proposition 1]{adamaszek2020homotopy} that the Vietoris--Rips filtration of a metric gluing is homotopy equivalent to the wedge sum of their Vietoris--Rips filtrations.
Explicitly, the natural inclusion
\[
\VR_r\cX \vee \VR_r\cY \to \VR_r(\cX \vee \cY)
\]
is a homotopy equivalence for each $r \in \R$.
From this, we deduce the following.

\medskip\corollary
Let $\cX$ and $\cY$ be pointed metric spaces.
For any linear cohomology operation $\theta$:
\begin{align*}
	\img_\theta^\VR(\cX \, \vee \, \cY) &\cong \, \img_\theta^\VR(\cX) \oplus \img_\theta^\VR(\cY), \\
	\ker_\theta^\VR(\cX \, \vee \, \cY) &\cong \, \ker_\theta^\VR(\cX) \oplus \ker_\theta^\VR(\cY).
\end{align*}

\subsubsection{}

Similarly, for any two $\R$-spaces $X$ and $Y$, the $\R$-space $X \times Y$, is defined by
\[
(X \times Y)_r = X_r \times Y_r \quad\text{and}\quad (X \times Y)_{s,t} = (X_{s,y} \times Y_{s,t}).
\]

For general cohomology operations there is no formula expressing its action on a product in terms of its action on the factors.
For Steenrod operations one does.

\medskip\theorem
Let $X$ and $Y$ be cellular $\R$-spaces.
For any $k \in \N$, the Steenrod persistence modules satisfy:
\begin{align*}
	\img_{\Sq^k}(X \times Y) &\cong\, \img\sum_{\mathclap{i+j=k}} \Sq^i_X \ot \Sq^j_Y\,, &
	\ker_{\Sq^k}(X \times Y) &\cong\, \ker\sum_{\mathclap{i+j=k}} \Sq^i_X \ot \Sq^j_Y\,, \\
	\img_{\rP^k}(X \times Y) &\cong\, \img\sum_{\mathclap{i+j=s}} \rP^i_X \otimes \rP^j_Y\,, &
	\ker_{\rP^k}(X \times Y) &\cong\, \ker\sum_{\mathclap{i+j=s}} \rP^i_X \otimes \rP^j_Y\,. &
\end{align*}

\begin{proof}
	For any \(r \in \R\), by the K\"unneth formula with field coefficients, the projections $X_r \leftarrow X_r \times Y_r \to Y_r$ induce an isomorphism $\rH^*(X_r) \otimes \rH^*(Y_r) \to \rH^*(X_r \times Y_r)$ natural with respect to \(r\).
	The claim now follows from the Cartan formula and the naturality of Steenrod operations.
\end{proof}

\subsubsection{}

Given two metric spaces $\cX$ and $\cY$.
The $\ell^\infty$ metric on $\cX \times \cY$ is defined by
\[
d(x \times y, \, x' \times y') = \max\set{d(x,x'), d(y,y')}.
\]
with this metric we have the following.

\medskip\corollary
For any two metric spaces $\cX$ and $\cY$,
Steenrod persistence modules satisfy:
\begin{align*}
	\img\Sq^k_{\VR\cX \times \VR\cY} &\cong\, \img\sum_{\mathclap{i+j=k}} \Sq^i_{\VR\cX} \ot \Sq^j_{\VR\cY}\,, \\ \ker\Sq^k_{\VR\cX \times \VR\cY} &\cong\, \ker\sum_{\mathclap{i+j=k}} \Sq^i_{\VR\cX} \ot \Sq^j_{\VR\cY}\,, \\
	\img\rP^k_{\VR\cX \times \VR\cY} &\cong\, \img\sum_{\mathclap{i+j=k}} \rP^i_{\VR\cX} \otimes \rP^j_{\VR\cY}\,, \\
	\ker\rP^k_{\VR\cX \times \VR\cY} &\cong\, \ker\sum_{\mathclap{i+j=k}} \rP^i_{\VR\cX} \otimes \rP^j_{\VR\cY}\,,
\end{align*}
for any $k \in \N$.

\begin{proof}
	This follows from the proof of \cite[Theorem~6.1(1)]{lim2020vietoris}, which constructs a natural isomorphism $\rH(\VR\cX) \otimes \rH(\VR\cY) \cong \rH(\VR(\cX \times \cY))$.\anibal{Note to self: Think more about this and fix notation}
\end{proof}